### 2018-02-06 ipsj/reply_letter/paper.tex
Add reply_letter

\documentclass{article}

\usepackage{a4wide}
\usepackage{graphicx}
\usepackage{fancybox}
\usepackage{ascmac}
\usepackage{xspace}
%\usepackage{pdfpages}
\usepackage{pifont}
\usepackage{array}

\newcommand\st{\textsuperscript{st}\xspace}
\newcommand\nd{\textsuperscript{nd}\xspace}
\newcommand\rd{\textsuperscript{rd}\xspace}

\begin{document}

% cover letter
\begin{flushleft}
  2018-03-03\newline 
  Journal of Information Processing\newline
  Submission ID: 18-Y008e\newline
  Title: Component-Based mruby Platform for IoT Devices\newline
  Authors: Takuro Yamamoto, Takuma Hara, Takuya Ishikawa, Hiroshi Oyama, Hiroaki Takada, and Takuya Azumi\newline
\end{flushleft}

\textbf{Dear Editor and Reviewer,}\newline

We highly appreciate the insightful suggestions and detailed valuable comments on our paper.
The suggestions of the reviewers are very helpful for us and the suggestions are now incorporated in the revised paper as follows.
We have attached the last paper review and revised paper in our reply letter.
In the revised paper, newly added and modified sentences are written in the bold and red colored font so that the reviewers can easily find them.
We hope the editor and the reviewers will be satisfied with our replies to the comments and the revised paper.
\newline\newline

\begin{flushleft}
  Yours sincerely,\newline

  Takuro Yamamoto\newline
  Osaka University\newline
  Email address: t-yamamoto@hopf.sys.es.osaka-u.ac.jp\newline
\end{flushleft}

\clearpage

\section{Response to 1\st reviewer}

\begin{enumerate}

\item \begin{flushleft}
    \textbf{Comment:}
  \end{flushleft}

  \begin{flushleft}
    \textbf{Our reply:}

\begin{itembox}[|]{}
\end{itembox}\\
  \end{flushleft}

\item \begin{flushleft}
    \textbf{Comment:}
  \end{flushleft}

  \begin{flushleft}
    \textbf{Our reply:}
  \end{flushleft}

\end{enumerate}

\section{Response to 2\nd reviewer}

\begin{enumerate}

\item \begin{flushleft}
    \textbf{Comment:}
  \end{flushleft}

  \begin{flushleft}
    \textbf{Our reply:}
  \end{flushleft}

\item \begin{flushleft}
    \textbf{Comment:}
  \end{flushleft}

  \begin{flushleft}
    \textbf{Our reply:}
  \end{flushleft}

\end{enumerate}

\end{document}