### 2018-02-24 ipsj/chosha.tex
Add choshasyoukai

\begin{biography}
\profile{Takuro Yamamoto}
{
received his B.E. degree from Osaka University in 2016.
His research interests include real-time and embedded systems\@.
}
\profile{Takuma Hara}
{
\@.
}
\profile{Takuya Ishikawa}
{
\@.
}
\profile{Hiroshi Oyama}
{
received Dr. Engineering degree from Gifu University in 2002.
He works at Okuma Corporation as Senior Engineer and also works at Nagoya University as Visiting Professor.
He is studying software components and programming language for embedded systems\@.
}
\profile{Hiroaki Takada}
{
is a professor at Institute of Innovation for Future Society, Nagoya University.
He is also a professor and the Executive Director of the Center for Embedded Computing Systems (NCES), the Graduate School of Informatics, Nagoya University.
He received his Ph.D. degree in Information Science from University of Tokyo in 1996.
He was a Research Associate at University of Tokyo from 1989 to 1997, and was a Lecturer and then an Associate Professor at Toyohashi University of Technology from 1998 to 2003.
His research interests include real-time operating systems, real-time scheduling theory, and embedded system design.
He is a member of ACM, IEEE, IEICE, JSSST, and JSAE\@.
}
\profile{Takuya Azumi}
{
is an Assistant Professor at the Graduate School of Engineering Science, Osaka University. 
He received his Ph.D. degree from the Graduate School of Information Science, Nagoya University.
From 2008 to 2010, he was under the research fellowship for young scientists for Japan Society for the Promotion of Science.
From 2010 to 2014, he was an Assistant Professor at the College of Information Science and Engineering, Ritsumeikan University.
His research interests include real-time operating systems and component based development.
He is a member of IEEE, ACM, IPSJ, and IEICE\@.
}

\end{biography}