### 2018-05-27 ipsj/chosha.tex
Final submit

\begin{biography}
\profile{Takuro Yamamoto}
{
received his B.E. degree from Osaka University in 2016.
His research interests include real-time and embedded systems\@.
}
\profile{Takuma Hara}
{
received his M.S. degree in Information Science from Nagoya University in 2013.
His research interests include embedded component technology\@.
}
\profile{Takuya Ishikawa}
{
received his Ph.D. degree in information science from Nagoya University in 2013.
He was a researcher and then a designated assistant professor at the Center for Embedded Computing Systems, Nagoya University from 2014 to 2016.
His research interests include real-time scheduling theory, real-time operating systems, and embedded software platform\@.
}

\profile{Hiroshi Oyama}
{
received his Dr. Engineering degree from Gifu University in 2002.
He works at Okuma Corporation as Senior Engineer and also works at Nagoya University as Visiting Professor.
He is studying software components and programming language for embedded systems\@.
}
\profile{Hiroaki Takada}
{
is a professor at Institutes of Innovation for Future Society, Nagoya University.
He is also a professor and the Executive Director of the Center for Embedded Computing Systems (NCES), the Graduate School of Informatics, Nagoya University.
He received his Ph.D. degree in Information Science from University of Tokyo in 1996.
He was a Research Associate at University of Tokyo from 1989 to 1997, and was a Lecturer and then an Associate Professor at Toyohashi University of Technology from 1998 to 2003.
His research interests include real-time operating systems, real-time scheduling theory, and embedded system design.
He is a member of ACM, IEEE, IEICE, JSSST, and JSAE\@.
}
\profile{Takuya Azumi}
{
is an Associate Professor at the Graduate School of Science and Engineering, Saitama University.
He received his Ph.D. degree from the Graduate School of Information Science, Nagoya University. 
From 2008 to 2010, he was under the research fellowship for young scientists for Japan Society for the Promotion of Science. 
From 2010 to 2014, he was an Assistant Professor at the College of Information Science and Engineering, Ritsumeikan University. 
From 2014 to 2018, he was an Assistant Professor at the Graduate School of Engineering Science, Osaka University. 
His research interests include real-time operating systems and component-based development.
He is a member of IEEE, ACM, IPSJ, and IEICE\@.
}

\end{biography}